Legal Services Market in Russia: Regional Specialization

### 2024-03-30 legal-companies-regional-specialization/Paper.Rmd
Minor improvements

Abstract: This paper focuses on the spatial distribution of the legal services market in Russia. The aim is to identify the most specialized regions (constituent entities) using a quantitative technique based on the relative numbers of employees. The research utilizes open data distributed by the Federal Tax Service of Russia and the Federal Statistics Service. The regional specialization is measured using two metrics: concentration and localization. High concentration indicates regions with legal services of national importance, while high localization marks regions where legal services have local importance. The study's results reveal several clusters of legal services. These clusters are likely linked to the high level of economic activity in the regions. The location of appeal and cassation courts may also partially explain the clusters.

@@

**Data.** This study utilizes the open data on Russian small and medium-sized businesses (SMBs) distributed by the Federal Tax Service (FTS)[^1]. Additionally, the FTS open data on the average number of employees of organizations[^2] was used. The author extracted information about legal firms (companies and individual entrepreneurs) that existed in 2021, including their region of registration and number of employees, from the source datasets. To filter out non-legal firms, the activity code was restricted to 69.10. The data was processed using the command-line Python tool *ru-smb-companies*[^3], developed by the author of this paper. In addition to the company's data, official statistical information from the Federal Statistics Service of Russia (Rosstat) was utilized to determine the average yearly employee count by region in 2021[^4].

@@

In certain studies, the localization metric is used to conclude about the specialization [@Romashina2019, @Bryantseva2022]. However, the authors of the Atlas propose a two-dimensional approach that combines concentration and specialization as follows. After calculating the metrics, the regions are ranked by localization, and the top 16 regions (approximately 20%) are considered to be areas where the legal services are of local importance. Next, the regions are ranked by concentration, and the cumulative concentration is calculated. Regions with a cumulative concentration below 0.8 are identified as areas where the legal services are of national importance, that is, these regions collectively account for 80% of the country's legal workers. Some regions are present in both two groups.

@@

**Results and discussion.** Figure 1 shows the tile-grid map of regional specialization in legal services. Several clusters are distinguishable based on the importance of legal services. These include (1) Moscow (with the surrounding Moscow oblast), (2) Saint Petersburg, (3) the group of southern regions (Krasnodar and Stavropol krai, Voronezh, Belgorod, and Rostov oblasti), (4) the Volga belt formed by Volgograd, Saratov, Samara oblasti, Perm krai, Tatarstan and Udmurt republics, (5) the Urals cluster with Sverdlovsk and Chelyabinsk oblasti as the core and Bashkir republic on the peripheries, (6) Siberian row (Tuymen, Omsk, Novosibirsk, Keverovo oblasti, Altai and Krasnoyarsk krai, Irkutsk oblast). Nizhniy Novgorod oblast and Primorsky krai can be treated as single-region clusters. Some regions, such as Kaliningrad, Tver, Ivanovo, Murmansk oblasti, and Kamchatka krai, have legal services of local importance that are not part of any cluster.

Further investigation is required to determine the reasons for the observed spatial patterns. However, legal services are likely concentrated in Moscow and Saint Petersburg due to their high levels of economic development and concentration. The other clusters can also be explained by their level of economic development. The same hypothesis may be true for the two single-region clusters. Most specialized regions have at least one ordinary or commercial appeals or cassation court. This concentration or localization of legal services in these regions may be partially explained by the presence of such courts, which act as “gravitational centers” for both legal services customers and providers. However, this hypothesis requires further testing.

>>>>

Abstract: This paper focuses on the spatial distribution of the legal services market in Russia. The aim is to identify the most specialized regions (constituent entities) using a quantitative technique based on the relative numbers of employees. The research utilizes open data distributed by the Federal Tax Service and the Federal Statistics Service of Russia. The regional specialization is measured using two metrics: concentration and localization. High concentration indicates regions with legal services of national importance, while high localization marks regions where legal services have local importance. The study's results reveal several clusters of legal services. These clusters are likely linked to the high level of economic activity in the regions. The location of appeal and cassation courts may also partially explain the clusters.

@@

**Data.** This study utilizes the open data on Russian small and medium-sized businesses (SMBs) distributed by the Federal Tax Service (FTS)[^1]. Additionally, the FTS open data on the average number of employees of organizations[^2] is used. The author extracted information about legal firms (companies and individual entrepreneurs) that existed in 2021, including their region of registration and number of employees, from the source datasets. To filter out non-legal firms, the main activity code was restricted to 69.10. The data was processed using the command-line Python tool *ru-smb-companies*[^3], developed by the author of this paper. In addition to the company's data, official statistical information from the Federal Statistics Service of Russia (Rosstat) was utilized to determine the average yearly employees count by region in 2021[^4].

@@

In certain studies, only the localization metric is used to conclude about the specialization [@Romashina2019, @Bryantseva2022]. However, the authors of the Atlas propose a two-dimensional approach that combines concentration and specialization as follows. After calculating the metrics, the regions are ranked by localization, and the top 16 regions (approximately 20%) are considered to be areas where the legal services are of local importance. Next, the regions are ranked by concentration, and the cumulative concentration is calculated. Regions with a cumulative concentration below 0.8 are identified as areas where the legal services are of national importance, that is, these regions collectively account for 80% of the country's legal workers. Some regions are present in both two groups.

@@

**Results and discussion.** Figure 1 shows the tile-grid map of regional specialization in legal services. Several clusters are distinguishable based on the importance of legal services. These include (1) Moscow (with the surrounding Moscow region), (2) Saint Petersburg, (3) the group of southern regions (Krasnodar and Stavropol territories, Voronezh, Belgorod, and Rostov regions), (4) the Volga belt formed by Volgograd, Saratov, Samara regions, Perm territory, Tatarstan and Udmurt republics, (5) the Urals cluster with Sverdlovsk and Chelyabinsk regions as the core and Bashkir republic on the peripheries, (6) Siberian row (Tyumen, Omsk, Novosibirsk, Kemerovo regions, Altai and Krasnoyarsk territories, Irkutsk region). Nizhniy Novgorod region and Primorsky territory can be treated as single-region clusters. Some regions, such as Kaliningrad, Tver, Ivanovo, Murmansk regions, and Kamchatka territory, have legal services of local importance and are not a part of any cluster.

Further investigation is required to determine the reasons for the observed spatial patterns. However, legal services are likely concentrated in Moscow and Saint Petersburg due to their high levels of economic development and concentration. The other clusters can also be explained by their level of economic development. The same hypothesis may be true for the two single-region clusters. Most specialized regions have at least one ordinary or commercial appeals or cassation court. The concentration or localization of legal services in these regions may be partially explained by the presence of such courts, which act as “gravitational centers” for both legal services customers and providers. However, this hypothesis requires further testing.
